Spatial Analysis

### 2021-09-28 Scripts/spatial_analisis.Rmd
Created Figure one and two based on new port analysis. (Top 74th per)

```{r setup, include=FALSE}
library(MyFunctions)
MyFunctions::my_lib(c("ggmap","sf","st","tidyverse","tools","readr","data.table","maps","viridis","wesanderson","knitr","kableExtra","plotly","ggrepel","ggsflabel","janitor","interp","rnaturalearth","readxl"))

@@

@@

# 
ports_data <- bind_rows(read_excel_allsheets(ports_path)) %>%
  clean_names() %>%
  mutate_if(is.character,
              ~ str_to_lower(.)) %>%
    filter(species_name %in% c("summer flounder","black sea bass","scup"))

@@

noaa_port <- port_sf %>% 
  filter(distance_from_shore_m == 0) %>% 
  mutate(label = str_to_lower(label)) %>% 
  filter(label %in% noaa_ports)

# Get number of ports per state
noaa_ports_n_data <- noaa_port %>% 
  as.data.frame() %>% 
    group_by(state) %>% 
    summarise(noaa_ports = n())

# Get number of ports in the NOAA database

n_ports_data <- ports_data %>% 
  group_by(landing_state) %>% 
  summarise(noaa_ports = length(unique(landing_port))) %>% 
  mutate(abrev = str_to_upper(landing_state)) %>% 
  left_join(land_sf) %>% 
  select(state,noaa_ports) %>% 
  filter(!is.na(state)) %>% 
  left_join(noaa_ports_n_data)

# plot it
gridExtra::grid.arrange(
  port_sf %>% 
    mutate(label = str_to_lower(label)) %>% 
    filter(distance_from_shore_m == 0,
           label %in% noaa_ports
           ) %>% 
    ggplot()  +
    geom_sf(aes(color = jurisdicti)) +
    scale_color_manual(values = state_pallet) +
    theme_dark(),
    # theme(legend.position = ""),
  tableGrob(n_ports_data),
  ncol = 2
)

@@

######4.2 Data from NOAA fisheries 
https://www.fisheries.noaa.gov/foss/f?p=215:11:15514950376034::NO:::

Delaware data was colected from NOAA's [Snapshots of Human Communities and Fisheries in the Northeast
](https://apps-nefsc.fisheries.noaa.gov/read/socialsci/communitySnapshots.php)

noaa_ports <- data.table(
  port =str_to_sentence(c(
    "POINT JUDITH, RI", #41.359907914553574,-71.48323250872187,
    "BEAUFORT, NC", #34.7138879654054,-76.66239630117921,
    "HAMPTON, VA", #36.98600366773327, -76.35038306859545,
    "PT. PLEASANT, NJ", #40.08351847675102,-73.99190094110679,
    "NEWPORT NEWS, VA", #37.00933919295183, -76.48565862613168,
    "BELFORD, NJ", # 40.43918784734024, -74.0763054899462
    "MONTAUK, NY", #41.07342262868038, -71.93585836088813
    "HOBUCKEN, NC", #34.94844567982384, -76.74333419560425
    "WANCHESE, NC", #35.83524984550334, -75.61121671722921
    "NEW BEDFORD, MA", # 41.630494052326306, -70.91580669134055
    "CAPE MAY, NJ", # 38.94979719126508, -74.90280944790463
    "ORIENTAL, NC", # 35.013571084373126, -76.66793114333704
    "CHINCOTEAGUE, VA", # 37.91788094070204, -75.35947580478219
    "ENGELHARD, NC", # 35.262929243118776, -76.08680452670508
    "STONINGTON, CT", # 41.327282204484035, -71.9223721359251
    "LONG BEACH/BARNEGAT LIGHT, NJ", #39.752855929900484, -74.11329310115902
    "OCEAN CITY, MD", #38.3802922004859, -75.05089736920318
    "Indian river, DE",
    "Portland, ME", #43.66344794728992, -70.22937554378885
    "Portsmouth, NH" #43.05334694203953, -70.7126760833714
  )),
  lat = c(
    41.359907914553574,
    34.7138879654054,
    36.98600366773327, 
    40.08351847675102,
    37.00933919295183, 
    40.43918784734024, 
    41.07342262868038,
    34.94844567982384, 
    35.83524984550334,
    41.630494052326306,
    38.94979719126508,
    35.013571084373126,
    37.91788094070204, 
    35.262929243118776, 
    41.327282204484035, 
    39.752855929900484, 
    38.3802922004859,
    38.62113593403746,
    43.66344794728992,
    43.05334694203953
  ),
  lon = c(
    -71.48323250872187,
    -76.66239630117921,
    -76.35038306859545,
    -73.99190094110679,
    -76.48565862613168,
    -74.0763054899462,
    -71.93585836088813,
    -76.74333419560425,
    -75.61121671722921,
    -70.91580669134055,
    -74.90280944790463,
    -76.66793114333704,
    -75.35947580478219,
    -76.08680452670508,
    -71.9223721359251,
    -74.11329310115902,
    -75.05089736920318,
    -75.06397021718345,
     -70.22937554378885,
    -70.7126760833714
  )
) %>% 
  separate(port, into = c("port","abrev"), sep = ',') %>% 
  mutate(abrev = str_squish(str_to_upper(abrev))) %>% 
  left_join(as.data.frame(land_sf)) %>% 
  left_join(noaa_ports) %>% 
  select(port,state = ID,lat,lon) %>% 
  mutate(state = str_to_sentence(state))

@@

# Visual exploration of all together (o.k! )
ggplot() +
  geom_sf(data = eez_sf, aes()) +
  geom_sf(data = land_sf, aes(fill = state), alpha = 0.3) +
  geom_sf(data = state_sf, aes(color = state), fill = NA) +
  geom_point(data = noaa_ports, aes(x = lon, y = lat)) +
  scale_color_manual(values = state_pallet) +
  scale_fill_manual(values = state_pallet) +
  theme_dark() +
  coord_sf(ylim = c(33,48))

@@

port_sf <- st_as_sf(noaa_ports,
                    coords =c("lon","lat"),
                    crs = 4326)

@@

ggplot() +
  geom_sf(data = eez_sf, aes(),fill = NA) +
  geom_sf(data = port_bf, aes(color = state), fill = NA) +
  scale_color_manual(values = state_pallet)+
  theme_dark()

@@

# ---------------- #
# Convert grid to sf 
# ---------------- #
grid_fp_sf <- st_as_sf(ne_grid,
             coords = c("lon", "lat"),
             crs = 4326) %>% 
  st_join(port_bf) %>% # join to port sf
  filter(!is.na(state)) %>% 
  distinct(index,state, .keep_all = TRUE) #removes like 400K overlapping rows

# Create data frame for future computations
# Note, will be used in next chunk
grid_fp_bf_dt <- as.data.frame(grid_fp_sf) %>%
    select(index,state)

@@

# Get final df with index
grid_eez_fp_df <- as.data.frame(grid_eez_sf_fp) %>%
  select(state,index) %>%
  left_join(ne_grid,
            by = "index")

@@

# Load grid of fishing ports
grid_eez_fp_sf <-  st_read(my_path("D","Spatial/grid_eez_fp_sf", name = "grid_eez_fp.shp")) %>%
  select(index,state,geometry) %>% 
  mutate(method = "fishing_port",
         state = str_to_sentence(state))%>% 
  left_join(state_order)

@@

p_fp <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_eez_fp_sf) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color = group, fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    scale_color_manual(values = state_pallet) +
    scale_fill_manual(values = state_pallet) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45)) +
    labs(x = "", y = "", title = "B) Fishing ports approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot(grid_eez_fp_sf) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    geom_sf(aes(color = group),size = 0.1, alpha = 0.3) +
    facet_wrap(~ group) +
    theme(legend.position = "top") +
    scale_color_manual(values = state_pallet,
                       # labels = unique(grid_eez_fp_sf$group)) +
                       labels = grid_eez_fp_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

@@

ggsave(filename = "buffer_figurec.jpg",
       plot = combined_plot,
       path = my_path("R","Figures"),
       width = 11,
       height = 11
)

@@

unique(historic_spp$spp)
unique(historic_spp$region)

@@

state_lat <- historic_spp %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(lat = order,
    order = letters[1:11])

@@

@@

state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  # left_join(periods,
  #           by = "year") %>% 
  group_by(state,label,region,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(label)) %>% 
  spread(spatial,mean_per) %>% 
  mutate(difference = (fp-sw)/((fp+sw)/2)*100,
         difference = ifelse(difference > 100,100,
                             ifelse(difference < -100,-100,difference)
         )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(spp = gsub(" ","\n",spp),
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         )
  )

# The plot
map_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Percentage difference", alpha = 0.8) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

@@

state_change <-historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  # left_join(periods,
            # by = "year") %>% 
  group_by(state,label,region,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(label)
         ) %>% 
  select(-region) %>% 
  spread(label,mean_per) %>% 
  mutate(change = (Today-Reference)/(Today)*100,
         change = ifelse(change > 100,100,
                         ifelse(change < -100,-100,change)
         )
  ) %>%
  select(-4,-5) %>% 
  spread(spatial,change) %>% 
  mutate(difference = (fp-sw)/((fp+sw)/2)*100,
         difference = ifelse(difference > 100,100,
                             ifelse(difference < -100,-100,difference)
         )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         )
  )

# The plot
prop_chng_map <- land_sf %>% 
  left_join(state_change,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Change in proportion today\n relative to 1973-1984\n(%)", alpha = 0.8,
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c(seq(-100,75,25),">100")
                              ) +
  facet_grid(spatial~ spp) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5)

ggsave(filename = "temp_proportion_chng.jpg",
         plot = prop_chng_map,
         path = my_path("R","Figures"),
         width = 10,
         height = 10
  )

>>>>

```{r setup, include=FALSE}
library(MyFunctions)
MyFunctions::my_lib(c("ggmap","sf","st","tidyverse","tools","readr","data.table","maps","viridis","wesanderson","knitr","kableExtra","plotly","ggrepel","ggsflabel","janitor","interp","rnaturalearth","readxl","cowplot"))

@@

###### 4.1.1 Ports coordinates

noaa_ports <- data.table(
  port =str_to_sentence(c(
    "POINT JUDITH, RI", #41.359907914553574,-71.48323250872187,
    "BEAUFORT, NC", #34.7138879654054,-76.66239630117921,
    "HAMPTON, VA", #36.98600366773327, -76.35038306859545,
    "POINT PLEASANT, NJ", #40.08351847675102,-73.99190094110679,
    "NEWPORT NEWS, VA", #37.00933919295183, -76.48565862613168,
    "BELFORD, NJ", # 40.43918784734024, -74.0763054899462
    "MONTAUK, NY", #41.07342262868038, -71.93585836088813
    "HOBUCKEN, NC", #34.94844567982384, -76.74333419560425
    "WANCHESE, NC", #35.83524984550334, -75.61121671722921
    "NEW BEDFORD, MA", # 41.630494052326306, -70.91580669134055
    "CAPE MAY, NJ", # 38.94979719126508, -74.90280944790463
    "ORIENTAL, NC", # 35.013571084373126, -76.66793114333704
    "CHINCOTEAGUE, VA", # 37.91788094070204, -75.35947580478219
    "ENGELHARD, NC", # 35.262929243118776, -76.08680452670508
    "STONINGTON, CT", # 41.327282204484035, -71.9223721359251
    "LONG BEACH/BARNEGAT LIGHT, NJ", #39.752855929900484, -74.11329310115902
    "OCEAN CITY harbor, MD", #38.3802922004859, -75.05089736920318
    # Manually included these ports as top State ports from NOAA
    "Indian river, DE",
    "Portland, ME", #43.66344794728992, -70.22937554378885
    "Portsmouth, NH", #43.05334694203953, -70.7126760833714
    # More ports
    "Chincoteague, VA", # 37.94295896302465, -75.38415668579887
    "sea isle city,NJ ", # 39.14912656323854, -74.68797202847871
    "lewes, DE, ", # 38.78943640990542, -75.06343692610851
    "virginia beach, VA" # 36.859343385722504, -75.94328398140065
  )),
  lat = c(
    41.359907914553574,
    34.7138879654054,
    36.98600366773327, 
    40.08351847675102,
    37.00933919295183, 
    40.43918784734024, 
    41.07342262868038,
    34.94844567982384, 
    35.83524984550334,
    41.630494052326306,
    38.94979719126508,
    35.013571084373126,
    37.91788094070204, 
    35.262929243118776, 
    41.327282204484035, 
    39.752855929900484, 
    38.3802922004859,
    38.62113593403746,
    43.66344794728992,
    43.05334694203953,
    37.94295896302465,
    39.14912656323854,
    38.78943640990542,
    36.859343385722504
  ),
  lon = c(
    -71.48323250872187,
    -76.66239630117921,
    -76.35038306859545,
    -73.99190094110679,
    -76.48565862613168,
    -74.0763054899462,
    -71.93585836088813,
    -76.74333419560425,
    -75.61121671722921,
    -70.91580669134055,
    -74.90280944790463,
    -76.66793114333704,
    -75.35947580478219,
    -76.08680452670508,
    -71.9223721359251,
    -74.11329310115902,
    -75.05089736920318,
    -75.06397021718345,
    -70.22937554378885,
    -70.7126760833714,
    -75.38415668579887,
    -74.68797202847871,
    -75.06343692610851,
    -75.94328398140065
  )
) %>% 
  separate(port, into = c("landing_port","landing_state"), sep = ',') %>% 
  mutate(landing_state = str_squish(str_to_upper(landing_state)),
         landing_port = str_to_lower(landing_port))

###### 4.1.1 Top ports

@@

# Load ports
ports_data <- bind_rows(read_excel_allsheets(ports_path)) %>%
  clean_names() %>%
  mutate_if(is.character,
              ~ str_to_lower(.)) %>%
    filter(species_name %in% c("summer flounder","black sea bass","scup"))

@@

# Get number of ports in the NOAA database
top10_ports <- ports_data %>% 
  mutate(landing_state = str_to_upper(landing_state)) %>% 
  filter(landing_state %in% c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")) %>% 
  group_by(year,landing_port,landing_state,species_name) %>% 
  summarise(
    year_ton = sum(landed_weight),
    year_value = sum(value),
    noaa_ports = length(unique(landing_port))
    ) %>% 
  group_by(landing_port,landing_state,species_name) %>% 
  summarise_if(is.numeric,sum) %>% 
  group_by(species_name) %>% 
  top_n(10,year_ton)

@@

###### 4.1.1 Visualization

top10_ports %>% 
  left_join(noaa_ports) %>% 
  ggplot() +
  geom_point(
    aes(
      x = lon,
      y = lat,
      shape = species_name, 
      color = species_name
    ),
    size = 5
  )

```{r}
# Get number of ports in the NOAA database
total_ports <- ports_data %>% 
  mutate(landing_state = str_to_upper(landing_state)) %>% 
  filter(landing_state %in% c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")) %>% 
  group_by(species_name) %>% 
  summarise(
    total_ton = sum(landed_weight, na.rm = T),
    total_value = sum(value, na.rm = T)
    )

# Per landings

per_ports <-
  ports_data %>% 
  mutate(landing_state = str_to_upper(landing_state)) %>% 
  filter(landing_state %in% c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")) %>% 
  group_by(landing_port,species_name) %>% 
  summarise(
    port_ton = sum(landed_weight),
    port_value = sum(value)
    ) %>% 
  left_join(total_ports,
            by = "species_name") %>% 
    mutate(
      per_tons = round((port_ton/total_ton)*100,2),
      per_value = round((port_value/total_value)*100,2)
    ) %>% 
    # select(landing_port,species_name,per_tons,everything()) %>% 
    ungroup() %>% 
    arrange(desc(per_tons)) %>% 
    group_by(species_name) %>% 
    mutate(cum_per = cumsum(per_tons)) %>% 
    filter(cum_per <= 75) %>% 
  select(landing_port,species_name,port_ton,per_tons,cum_per,everything()) %>% 
  arrange(desc(species_name,cum_per))

per_ports

per_ports %>% 
  left_join(noaa_ports) %>% 
  # Double ceck coordinates are all there
  # filter(is.na(lon))
  ggplot() +
  geom_point(
    aes(
      x = lon,
      y = lat,
      shape = species_name,
      color = species_name
    ),
    size = 5
  )

ports_df <-  per_ports %>% 
  left_join(noaa_ports)

@@

# Visual exploration of all together (o.k! )
ggplot() +
  geom_sf(data = eez_sf, aes()) +
  geom_sf(data = land_sf, aes(fill = state), alpha = 0.3) +
  geom_sf(data = state_sf, aes(color = state), fill = NA) +
  geom_point(data = ports_df, aes(x = lon, y = lat,
                                  shape = species_name), size = 3, alpha = 0.7) +
  scale_color_manual(values = state_pallet) +
  scale_fill_manual(values = state_pallet) +
  theme_dark() +
  coord_sf(ylim = c(33,48))

@@

port_sf <- st_as_sf(ports_df,
                    coords =c("lon","lat"),
                    crs = 4326)

@@

ggplot() +
  geom_sf(data = eez_sf, aes(),fill = NA) +
  geom_sf(data = port_bf, aes(color = species_name), fill = NA) +
  scale_color_manual(values = state_pallet)+
  theme_dark()

@@

# ---------------- #
# Convert grid to sf 
# ---------------- #
grid_fp_sf <- st_as_sf(ne_grid,
             coords = c("lon", "lat"),
             crs = 4326) %>% 
  st_join(port_bf) %>% # join to port sf
  filter(!is.na(landing_port)) %>%
  distinct(index,landing_port, .keep_all = TRUE) #removes like 400K overlapping rows

# Create data frame for future computations
# Note, will be used in next chunk
grid_fp_bf_dt <- as.data.frame(grid_fp_sf) %>%
    select(index,landing_port,species_name)

@@

# Get final df with index
grid_eez_fp_df <- as.data.frame(grid_eez_sf_fp) %>%
  select(index,landing_port,species_name) %>% 
  left_join(ne_grid,
            by = "index")

@@

# Load grid of fishing ports
grid_eez_fp_sf <-  st_read(my_path("D","Spatial/grid_eez_fp_sf", name = "grid_eez_fp.shp")) %>%
  select(index,abrev=lndng_s,geometry) %>% 
  mutate(method = "fishing_port"#,
         # state = str_to_sentence(landing_port)
         )%>% 
  left_join(state_order)

@@

grid_eez_fp_sf <-
  grid_eez_sw_sf %>% 
  filter(!group %in% grid_eez_fp_sf$group) %>% 
  mutate(index = NA) %>% 
  bind_rows(grid_eez_fp_sf)

p_fp <- gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(data = subset(grid_eez_fp_sf, index != "NA")) +
    geom_sf(data = eez_sf, aes(), fill = "white") + 
    geom_sf(aes(color = group, fill = abrev), alpha = 0.3) +
    geom_sf(data = land_sf, aes()) +
    ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
                                   size = 4,
                                   box.padding = 0.10,
                                   hjust = 1) +
    # scale_color_manual(values = state_pallet) +
    # scale_fill_manual(values = state_pallet) +
    scale_color_manual(values = c(state_pallet[3:4],state_pallet[6:13])) +
    scale_fill_manual(values = c(state_pallet[3:4],state_pallet[6:13])) +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 13) +
    coord_sf(ylim = c(30,48)) +
    scale_y_continuous(breaks = c(30,35,40,45)) +
    labs(x = "", y = "", title = "B) Fishing ports approach") +
    theme(plot.title = element_text(size = 20)),
  # Showing each state separately
  ggplot() +
    geom_sf(data = subset(grid_eez_fp_sf, index = "NA"), color = "white") +
    geom_sf(data = subset(grid_eez_fp_sf, index != "NA"), aes(color = group),size = 0.1, alpha = 0.3) +
    geom_sf(data = land_sf, aes(), fill = "grey80") +
    facet_wrap(~ group) +
    geom_point(data = rename(ports_df, group = landing_state), aes(x = lon, y = lat), color = "#E6A0C4") +
    theme(legend.position = "top") +
    # scale_color_manual(values = state_pallet,
    scale_color_manual(values = c(state_pallet[3:4],state_pallet[6:13]),
                       # labels = unique(grid_eez_fp_sf$group)) +
                       labels = grid_eez_fp_sf %>% arrange(order) %>%  pull(state) %>% unique()) +
    ggtitle("") +
    my_ggtheme_p(leg_pos = "",
                 ax_tx_s = 11,
                 axx_tx_ang = 45,
                 hjust = 1
    ),
  nrow = 1)

@@

ggsave(filename = "buffer_figure.jpg",
       plot = combined_plot,
       path = my_path("R","Figures"),
       width = 11,
       height = 11
       )

@@

unique(historic_spp$spp)
unique(historic_spp$region)
unique(historic_spp$region)

@@

state_lat <- historic_spp %>%
  group_by(state) %>% 
  summarise(
    order = mean(lat)
  ) %>% 
  filter(!is.na(state)) %>% 
  mutate(abrev = c("CT","DE","ME","MD","MA","NH","NJ","NY","NC","RI","VA")
  ) %>% 
  arrange(desc(order)) %>% 
  mutate(lat = order,
    order = letters[1:11])

@@

# Interpolation result

# Interpolation data
interpol_result <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label),
         region == "Northeast US Spring") %>% 
  group_by(label,index,spp,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue))

# --------------------- #
# Biomass Centroid shift
# --------------------- #

# Estimate total interpolation
total_inter <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label),
         region == "Northeast US Spring") %>% 
  group_by(spp,index) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  group_by(spp) %>% 
  summarise(total_cpue = sum(mean_cpue, na.rm = T))

# Estimate centroid based on top 10 grids with higer abundance
centroid_db <- historic_spp %>% 
  mutate(
    label = ifelse(year >= 1980 & year <= 2001,"Reference",
                   ifelse(year > 2001,"Today",NA)
    )
  ) %>% 
  filter(!is.na(label),
         region == "Northeast US Spring") %>% 
  group_by(label,index,spp,lat,lon) %>% 
  summarise(mean_cpue = mean(value,na.rm=T)) %>% 
  filter(!is.na(mean_cpue)) %>% 
  left_join(total_inter) %>% 
  mutate(
    per_cpue = round((mean_cpue/total_cpue)*100,2)
  ) %>% 
  group_by(spp,label) %>% 
  top_n(10,per_cpue) %>% 
  group_by(spp,label) %>% 
  summarise(
    lat = mean(lat),
    lon = mean(lon)
  )

inter_map <- ggplot() +
   geom_tile(data = subset(interpol_result, mean_cpue == 0),
            aes(
              x = lon,
              y = lat
            ),
            fill = "grey90",
              color = "grey90"
  ) + 
  geom_tile(data = subset(interpol_result, mean_cpue > 0),
            aes(
              x = lon,
              y = lat,
              fill = mean_cpue,
              color = mean_cpue
            )
  ) + 
  geom_point(data = centroid_db,
            aes(
              x = lon,
              y = lat
            ),
            shape = 3,
            color = "white"
  ) + 
  geom_sf(data = land_sf,aes()) +
  labs(x = "Longitude",
         y = "Latitude") +
  # ggsflabel::geom_sf_label_repel(data = land_sf, aes(label= abrev), 
  #                                  size = 4,
  #                                  box.padding = 0.10,
  #                                  hjust = 1) +
  scale_fill_viridis("Mean Interpolation", alpha = 0.8,
                     limits = c(0,40)) +
  scale_color_viridis("Mean Interpolation", alpha = 0.8, 
                      limits = c(0,40)) +
  facet_grid(label ~ spp) +
  my_ggtheme_m(map_type = "na", leg_pos = "bottom",ax_tx_s = 10,ax_tl_s = 12,leg_tx_s = 12) +
  theme(legend.key.width = unit(3,"line"))

ggsave("interpolation_fig.jpg",
       inter_map,
       path = my_path("R","Figures"),
       width = 8,
       height = 6)

@@

state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  # left_join(periods,
  #           by = "year") %>% 
  group_by(state,label,region,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(label)) %>% 
  spread(spatial,mean_per) %>% 
  mutate(
    difference = (fp-sw)/((fp+sw)/2)*100,
    # difference = fp-sw,
         difference = ifelse(difference > 100,100,
                             ifelse(difference < -100,-100,difference)
         )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(spp = gsub(" ","\n",spp),
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State waters","Difference")
         )
  )

# The plot
map_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8, 
                              breaks = seq(0,15,5),
                              limits=(c(0,15))
                              )+
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"))

diff_plot <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  filter(spatial == "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Percentage difference", alpha = 0.8,option = "C",
                              breaks = seq(-20,70,10),
                              limits=(c(-20,70))
                              ) +
  facet_grid(spatial~ spp+label) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        strip.background = element_blank(),
        strip.text.x = element_blank()
        )

@@

state_change <-historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100,
         label = ifelse(year >= 1980 & year <= 2001,"Reference",
                         ifelse(year > 2001,"Today",NA)
                         )
         ) %>% 
  group_by(state,label,region,spp,spatial) %>% 
  summarise(mean_per = round(mean(percentage)),.groups = "drop") %>% 
  # Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(label)
         ) %>% 
  select(-region) %>% 
  spread(label,mean_per) %>% 
  mutate(change = (Today-Reference)/(Today)*100,
         change = ifelse(change > 100,100,
                         ifelse(change < -100,-100,change)
         )
  ) %>%
  select(-4,-5) %>% 
  spread(spatial,change) %>% 
  mutate(difference = (fp-sw)/((fp+sw)/2)*100,
         difference = ifelse(difference > 100,100,
                             ifelse(difference < -100,-100,difference)
         )
  ) %>% 
  gather("spatial","mean_per",fp:difference) %>% 
  mutate(
         spatial = ifelse(spatial == "fp","Fishing ports",
                          ifelse(spatial == "sw","State Waters","Difference")
         ),
         mean_per= replace_na(mean_per,0)# for those 0/0
  )

# The plot
prop_chng_map <- land_sf %>% 
  left_join(state_change,
            by = "state") %>% 
  filter(spatial != "Difference") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Change in proportion today\nrelative to 1980-2001\n(%)", alpha = 0.8,
                              limits = c(-100,100),
                              breaks = seq(-100,100,25),
                              labels = c("<-100",seq(-75,75,25),">100")
                              ) +
  facet_grid(spatial~ spp) +
  labs(x = "", 
       y = "") +
  my_ggtheme_p(facet_tx_s = 20,
               leg_pos = "bottom",
               axx_tx_ang = 45,
               ax_tx_s = 12,
               ax_tl_s = 18,
               hjust = 1) +
  theme(legend.key.width = unit(4,"line"),
        legend.title.align=0.5)

# ggsave(filename = "temp_proportion_chng.jpg",
#          plot = prop_chng_map,
#          path = my_path("R","Figures"),
#          width = 10,
#          height = 10
#   )
